### 2025-06-13 task_progress/tables/6.4.4tables/tableF_hyper_medians.tex

\begin{tabular}{@{}p{3.2cm}*{8}{c}@{}}
      \toprule
      \textbf{超参数} & \textbf{Mode} & \textbf{Baran} & \textbf{Holoclean} & \textbf{BigDansing} & \textbf{BoostClean} & \textbf{Horizon} & \textbf{Scared} & \textbf{Unified} \\
      \midrule
      \multicolumn{9}{@{}l}{\textbf{Centroid型}}\\
\quad k & 30 ± 22 & 24 ± 21 & 25 ± 22.5 & 18 ± 19.2 & 18.5 ± 18 & 16 ± 13.2 & 15 ± 14.2 & 12 ± 16\
\quad n_components & 23.5 ± 17.2 & 14.5 ± 13 & 13.5 ± 15.5 & 12 ± 10 & 14.5 ± 16.5 & 11.5 ± 7.25 & 10.5 ± 10.5 & 9.5 ± 12\
\quad cov_type & tied (70%) & tied (63%) & tied (87%) & tied (60%) & tied (60%) & tied (52%) & tied (67%) & tied (53%)\
\multicolumn{9}{@{}l}{\textbf{Density型}}\\
\quad eps & 1.88 ± 0.63 & 1.99 ± 0.102 & 1.87 ± 0.302 & 1.9 ± 0.617 & 1.98 ± 0.145 & 1.94 ± 0.195 & 1.96 ± 0.175 & 1.96 ± 0.155\
\quad minPts & 5 ± 0 & 5 ± 2.25 & 5 ± 2 & 5 ± 1 & 5 ± 2.75 & 5 ± 2.25 & 5 ± 2 & 6 ± 5\
\multicolumn{9}{@{}l}{\textbf{Hierarchy型}}\\
\quad k & 6 ± 12.8 & 6 ± 7 & 5 ± 2.25 & 7 ± 8.25 & 5 ± 0.25 & 6 ± 3.25 & 5 ± 1.25 & 5 ± 3\
\quad linkage & single (95%) & single (93%) & single (100%) & single (87%) & single (90%) & single (78%) & single (95%) & single (90%)\
\quad metric & euclidean (50%) & manhattan (58%) & euclidean (58%) & euclidean (53%) & euclidean (52%) & manhattan (58%) & euclidean (58%) & euclidean (60%)\
      \bottomrule
    \end{tabular}